---
title: "Explorer Nodes"
author: "David Barkemeyer"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Explorer Nodes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `explorer` module takes a tree of objects of class `ExplorerNode` and visualizes
it using a `DT::datatable`. In this vignette we examine the structure of the tree
as a whole and of individual nodes. For simplicity we hereafter refer to objects of 
class `ExplorerNode` as nodes.

The explorer tree consists of nodes. Every explorer tree starts with a root node. 
Because `ExplorerNode` is an R6-Class we must call `ExplorerNode$new()` to instantiate 
a new node:

```{r, eval=FALSE}
tree <- ExplorerNode$new(
  id = "root",
  parent = NULL,
  is_group_node = TRUE
)
```

Every node has a unique id. 

<!--
You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.
-->
